# 00. Prepare Data

## Load Packages

```{r}
library(dplyr)
library(knitr)
library(readxl)
library(readr)
library(haven)
library(visdat)
library(sjstats)
library(openxlsx)
library(stargazer)
library(broom)
library(ggplot2)
library(sjPlot)
library(stringr)
library(caret)
library(future)
library(purrr)
library(furrr)
library(here)
library(svrep)
library(patchwork)
```

## Load Data (single level)
In the current setting, we view it as the whole population, and the original survey design would not be considered.
```{r}
nhanes_df = readRDS('data/pa_df_persub.rds') |> 
  dplyr::rename(Age = age_in_years_at_screening) |>
  dplyr::rename(weight = full_sample_2_year_mec_exam_weight)

## subjects in the sample
length(unique(nhanes_df$SEQN)) # N = 13491
head(nhanes_df)
```

## Select population samples

For now, we didn't set any variable restrictions (like Age and Sex)

## Estimate True Population Effect

$$
y_i(s) = \beta_0(s) + X_i \beta_1(s) + \epsilon_i(s), \quad s = 1, 2, \ldots, 1440
$$

-   $y_i(s)$: PA monitoring data for individual i

-   $X_i$: Scalar covariate, and we consider age here

-   $\beta_0(s)$: Time-varying regression coefficients

-   $\epsilon_i(s) \sim N(0, \sigma^2(s))$

We need find a way to estimate true coefficients, $\beta_{1(s)}$ and intercepts $\beta_{0(s)}$. Here, we consider the same way in fast univariate inference (FUI) to estimate point-wise effect. i.e. use generalized survey-weighted linear mixed effects model in each time point throughout the functional domain.

In the current scenario, we consider the simpler situation that single level and survey weighted generalized linear regression (svyglm).

```{r}
# Estimate point-wise effect
## Response Variable
min_columns <- nhanes_df %>% select(starts_with("min")) %>% names()
```

### Linear regression model

```{r}
## Obtain True Effects: Regression at each time point
# results.lr <- map_dfr(min_columns, ~ {
#   model <- lm(reformulate("Age", response = .x), data = nhanes_df)
#   tidy(model) %>%
#     filter(term == "Age") %>%
#     mutate(Variable = .x, Intercept = coef(model)[1], Beta = estimate) %>%
#     select(Variable, Intercept, Beta)
# })
# write_rds(results.lr, 'data/nhanes_true_effect_lr.rds')


## Smooth estimated effects
results.lr = read_rds('data/nhanes_true_effect_lr.rds')
intercept_smooth <- smooth.spline(results.lr$Intercept)
beta_smooth <- smooth.spline(results.lr$Beta)
  # Here, we use GCV to select spar (cubic smoothing spline)
  # what method we should use for smoothing true estimates?
results.lr$Intercept_smoothed <- predict(intercept_smooth)$y
results.lr$Beta_smoothed <- predict(beta_smooth)$y


## True Coefficients and intercept (& Visualization)
results.lr |>
  mutate(Minutes = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable) |>
  tidyr::pivot_longer(cols = c(Beta, Intercept_smoothed, Intercept, Beta_smoothed), 
                      names_to = "Type",
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value)) +
  geom_point(color = "blue",
             alpha = 0.6, size = 0.8) +
  geom_smooth(method = "loess",
              linewidth = 0.8, 
              color = "red", se = TRUE, span = 0.3) +
  facet_wrap(~ Type, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Beta") +
  theme(text = element_text(size = 14))


## Export smoothed estimates
results.lr = results.lr |>
  select(-Intercept, -Beta) |>
  rename(Intercept = Intercept_smoothed,
         Beta = Beta_smoothed)
# write_rds(results.lr, 'data/nhanes_true_effect_lr_smoothed.rds')
```

### Survey weighted linear regression model

**WTINT2YR:** Full Sample 2 Year Interview Weight (full sample of participants who completed the NHANES household interview over a two-year survey period)

-   All participants who completed the household interview

**WTMEC2YR:** Full Sample 2 Year MEC Exam Weight (full sample of participants who completed both the household interview and the Mobile Examination Center (MEC) examination over a two-year survey period)

-   Only participants who completed the MEC examination (Lab Measurements)

```{r}
library(survey)
## Regression at each time point
head(nhanes_df)
design <- svydesign(ids = ~1, 
                    weights = ~weight, 
                    strata = ~strata,
                    psu = ~psu,
                    data = nhanes_df)

# results.svylr <- map_dfr(min_columns, ~ {
#   model <- svyglm(reformulate("Age", response = .x),
#                   design = design)
#   tidy(model) %>%
#     filter(term == "Age") %>%
#     mutate(Variable = .x, Intercept = coef(model)[1], Beta = estimate) %>%
#     select(Variable, Intercept, Beta)
# })
results.svylr = read_rds('data/nhanes_true_effect_svylr.rds')

## True Coefficients and intercept (& Visualization)
results.svylr |>
  mutate(Minutes = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable) |>
  tidyr::pivot_longer(cols = c(Beta, Intercept), 
                      names_to = "Type",
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value)) +
  geom_point(color = "blue",
             alpha = 0.6, size = 0.8) +       # 绘制数据点
  geom_smooth(method = "loess",
              linewidth = 0.8, 
              color = "red", se = TRUE, span = 0.3) +
  facet_wrap(~ Type, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Beta") +
  theme(text = element_text(size = 14))
```

### Save true effects

```{r}
# write_rds(results.svylr, 'data/nhanes_true_effect_svylr.rds')
# write_rds(results.lr, 'data/nhanes_true_effect_lr.rds')
```

#### Compare true effects from svylm and lm

```{r}
results.svylr = read_rds('data/nhanes_true_effect_svylr.rds')
results.lr = read_rds('data/nhanes_true_effect_lr.rds')

df = merge(results.svylr, 
      results.lr, 
      by = 'Variable',
      suffixes = c("_svylr","_lr")) |>
  mutate(Minute = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable)

df |>
  select(Intercept_svylr, Beta_svylr, Minute) |>
  mutate(type = "svylr") |>
  rename(Intercept = Intercept_svylr, Beta = Beta_svylr) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minute) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minute, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

```{r}
source('src/00_sim_data_function.R')
source('src/01_sim_functions.R')
```

# 01. Non-Informative Sampling
## Ignore original survey setting

1.  Generate strata selection probabilities ($H=30$) from Dirichlet distribution
2.  Sample $N=\{10, 20, 30\}$ strata in the simulation
3.  For each strata $h$ in the selected strata:
    -   Standardize Age into mean=0 and SD=1
    -   Calculate inclusion probabilities $p_i = exp(Age_{i}/2)$
    -   Normalize: $p_{i'} = \frac{p_i}{\sum_{i \in h}p_i}$
    -   Adjust by expected number of individuals sampled $p_{i''} = min(p_{i'} * I_n * 2, 1)$
    -   Obtain sample using binomial sampling

### Simulate Data
```{r}
results.lr.tmp = read_rds('data/nhanes_true_effect_lr_smoothed.rds') |>
  mutate(Minutes = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable)

X_des = cbind(intercept = 0, nhanes_df |> select(Age))
num_strata = 30 # total strata
L = 150 # dimension of the functional domain
sampled_strata = 15
Y_obs = nhanes_df[,which(grepl('min_', colnames(nhanes_df)))[1:L] ]

sim.data = simulate_nhanes_survey_fosr(X_des, 
                                       Y_obs,
                                       L = L,
                                       sampled_strata = 15)
```

### Survey design/weighted method
#### Visualize Point-Estimate
```{r}
betaTilde = get_betatilde(sim.data,
              family = "gaussian", type = "design",
              model_formula = as.formula('Y~X'))
betaHat = get_betahat(betaTilde, L = L)

## Visualize differences
df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```


#### Visualize 95%CI
```{r}
betaTilde_boot = run_boots(
  sim.data,
  boot_type = 'Preston',
  betaHat = betaHat,
  L = L,
  num_boots = 300,
  model_formula = as.formula('Y~X')
)

mod_output = get_cis(betaTilde_boot,
        betaHat,
        smooth_for_ci = TRUE,
        smooth_for_variance = FALSE, # False
        L = L)

results = get_coverage_stats(mod_output,
                             beta_true = t(results.lr |> select(-Variable))[,1:L],
                             name = 'Preston',
                             # boot_type
                             L = L)

write_rds(betaHat, 'tmp_result/betaHat_NI_noOrigial_survey_betaHat.rds')
write_rds(betaTilde_boot, 'tmp_result/betaHat_NI_noOrigial_survey_betaTilde_boot.rds')
write_rds(mod_output, 'tmp_result/betaHat_NI_noOrigial_survey_mod_output.rds')
```




### No survey design/Unweighted
#### Visualize Point-Estimate
```{r}
betaTilde = get_betatilde(sim.data,
              family = "gaussian", type = "none",
              model_formula = as.formula('Y~X'))
betaHat = get_betahat(betaTilde, L = L)

df = as.data.frame(t(betaTilde)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

#### Visualize 95%CI
```{r}
betaTilde_boot = run_boots(
  sim.data,
  boot_type = 'Preston',
  betaHat = betaHat,
  L = L,
  num_boots = 300,
  model_formula = as.formula('Y~X')
)

mod_output = get_cis(betaTilde_boot,
        betaHat,
        smooth_for_ci = TRUE,
        smooth_for_variance = FALSE, # False
        L = L)

results = get_coverage_stats(mod_output,
                             beta_true = t(results.lr |> select(-Variable))[,1:L],
                             name = 'Preston',
                             # boot_type
                             L = L)

write_rds(betaHat, 'tmp_result/betaHat_NI_noOrigial_NOsurvey_betaHat.rds')
write_rds(betaTilde_boot, 'tmp_result/betaHat_NI_noOrigial_NOsurvey_betaTilde_boot.rds')
write_rds(mod_output, 'tmp_result/betaHat_NI_noOrigial_NOsurvey_mod_output.rds')
```



### Simulation results (N=200) 
#### Merge results
```{r}
L = 1440
all_results <- list.files("results/NI/NI_noWeight_FPCA_noSmooth_adjustSTD/", 
                          pattern = "NI_noWeight_FPCA_noSmooth_adjustSTD.*\\.rds", 
                          full.names = TRUE) %>%
  map_dfr(readRDS)

summary_design = all_results %>%
  group_by(boot_type, var) %>%
  summarise(
    MISE = mean(MISE),
    mean_joint_se = mean(mean_joint_se),
    mean_pw_se = mean(mean_pw_se),
    cover_joint_ratio = mean(cover_joint),
    cover_pw_ratio = mean(cover_pw) / L,
    .groups = 'drop'
  )

formatted_table <- summary_design %>%
  mutate(across(where(is.numeric), ~round(., 4)))  # 四舍五入到4位小数

kable(summary_design,
      digits = 4,  # control digital number 
      caption = "Summary Statistics",  # Simple Caption
      format = "html")

## In the current setting, the length of domain will affect joint-wise coverage ratio.
## How about changing knot points?
## Test it.
```

## Consider original weights
### Simulate data
```{r}
X_des = cbind(intercept = 0, nhanes_df |> select(Age, weight))
X_des$weight = scale(X_des$weight)
num_strata = 30 # total strata
L = 150 # dimension of the functional domain
sampled_strata = 15
Y_obs = nhanes_df[,which(grepl('min_', colnames(nhanes_df)))[1:L] ]

sim.data = simulate_nhanes_survey_fosr_w(X_des, Y_obs,
                                         L = L,
                                         sampled_strata = 15)
```

### Survey design/weighted method
#### Visualize Point-Estimate
```{r}
betaTilde = get_betatilde(sim.data,
              family = "gaussian", type = "design",
              model_formula = as.formula('Y~X'))
betaHat = get_betahat(betaTilde, L = L)

## Visualize differences
df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) + 
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

#### Visualize 95%CI
```{r}
betaTilde_boot = run_boots(
  sim.data,
  boot_type = 'Preston',
  betaHat = betaHat,
  L = L,
  num_boots = 300,
  model_formula = as.formula('Y~X')
)

mod_output = get_cis(betaTilde_boot,
        betaHat,
        smooth_for_ci = TRUE,
        smooth_for_variance = FALSE, # False
        L = L)

results = get_coverage_stats(mod_output,
                             beta_true = t(results.lr |> select(-Variable))[,1:L],
                             name = 'Preston',
                             # boot_type
                             L = L)

results

write_rds(betaHat, 'tmp_result/betaHat_NI_oW_survey_betaHat.rds')
write_rds(betaTilde_boot, 'tmp_result/betaHat_NI_oW_survey_betaTilde_boot.rds')
write_rds(mod_output, 'tmp_result/betaHat_NI_oW_survey_mod_output.rds')
```

#### Simulation multiple times
```{r}
nsim = 50
parallel = TRUE
ncores = parallelly::availableCores() - 1
L = 150
B = 300
family = "gaussian"
# boot_types = c('Rao-Wu-Yue-Beaumont', 'BRR', 'Preston', 'weighted', 'none')
# fit_types = c('design', 'design', 'design', 'weighted', 'none')

boot_types = c('Preston', 'weighed', 'none')
fit_types = c('design', 'weighted', 'none')

sim_res <- list() ## store simulation results
for (iter in 1:nsim) {
  set.seed(iter)
  x = try({
    data <- simulate_nhanes_survey_fosr_w(X_des, Y_obs, L = L, sampled_strata = 15)
    
    betaTilde = map(.x = fit_types, .f = get_betatilde, data = data, family = family)
    betaHat = map(.x = betaTilde, get_betahat, L = L)
    if (parallel) {
      plan(multisession, workers = ncores)
      res = future_map2(
        .x = boot_types,
        .y = betaHat,
        .f = run_boots,
        data = data,
        # family = family,
        num_boots = B,
        set_seed = TRUE,
        seed = 2025,
        L = L,
        .options = furrr_options(seed = TRUE)
      )
      cis = future_map2(
        .x = res,
        .y = betaHat,
        .f = get_cis,
        L = L,
        smooth_for_ci = FALSE,
        .options = furrr_options(seed = TRUE)
      )
      plan(sequential)
    } else{
      res = map2(
        .x = boot_types,
        .y = betaHat,
        .f = run_boots,
        data = data,
        # family = family,
        num_boots = B,
        set_seed = TRUE,
        seed = 2025,
        L = L)
      cis = map2(
        .x = res,
        .y = betaHat,
        .f = get_cis,
        smooth_for_ci = TRUE)
    }
    stats = map2(
      .x = cis,
      .y = boot_types,
      .f = get_coverage_stats,
      L = L,
      beta_true = t(results.lr |> select(-Variable))[,1:L]
    ) %>%
      list_rbind() %>%
      mutate(n = nrow(data), n_boot = B)
    sim_res[[iter]] <- stats
  })
  rm(x)
  print(iter)
}

result =
  sim_res %>%
  keep(., is.data.frame) %>%
  list_rbind(names_to = "id")

write_rds(result, 'tmp_result/betaHat_NI_oW_survey_results.rds')
```


### No survey design/Unweighted
#### Visualize Point-Estimate
```{r}
betaTilde = get_betatilde(sim.data,
              family = "gaussian", type = "none",
              model_formula = as.formula('Y~X'))
betaHat = get_betahat(betaTilde, L = L)

## Visualize differences
df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) + 
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

#### Visualize 95%CI
```{r}
betaTilde_boot = run_boots(
  sim.data,
  boot_type = 'Preston',
  betaHat = betaHat,
  L = L,
  num_boots = 300,
  model_formula = as.formula('Y~X')
)

mod_output = get_cis(betaTilde_boot,
        betaHat,
        smooth_for_ci = TRUE,
        smooth_for_variance = FALSE, # False
        L = L)

results = get_coverage_stats(mod_output,
                             beta_true = t(results.lr |> select(-Variable))[,1:L],
                             name = 'Preston', # boot_type
                             L = L)

write_rds(betaHat, 'tmp_result/betaHat_NI_oW_NOsurvey_betaHat.rds')
write_rds(betaTilde_boot, 'tmp_result/betaHat_NI_oW_NOsurvey_betaTilde_boot.rds')
write_rds(mod_output, 'tmp_result/betaHat_NI_oW_NOsurvey_mod_output.rds')
```

### Simulation results (N=200) 
#### Merge results
```{r}
L = 1440
all_results <- list.files("results/NI/NI_noWeight_noSmooth//", 
                          pattern = "NI_noWeight_noSmooth_.*\\.rds", 
                          full.names = TRUE) %>%
  map_dfr(readRDS)


summary_design = all_results %>%
  group_by(boot_type, var) %>%
  summarise(
    MISE = mean(MISE),
    mean_joint_se = mean(mean_joint_se),
    mean_pw_se = mean(mean_pw_se),
    cover_joint_ratio = mean(cover_joint),
    cover_pw_ratio = mean(cover_pw) / L,
    .groups = 'drop'
  )

formatted_table <- summary_design %>%
  mutate(across(where(is.numeric), ~round(., 4)))  # 四舍五入到4位小数

kable(summary_design,
      digits = 4,  # 控制小数位数
      caption = "Summary Statistics",  # 简单标题
      format = "html")

## 0422: increase knots for estimating covariance and smooth boostratped trajectory
## 0423: increase knots for fPCA, no use
## double: double the number of knots
## try different simulated number (1-200 ->1-400): 0.9075

## candidate steps: increase scaling factors
## knot doesn't matter after reaching certain values.
## visualize current issue
## try 


all_results |>
  filter(var == 'Beta') |>
  filter(boot_type == 'Preston') |>
  filter(cover_joint == FALSE)
```

### [DROP] Consider orginal strata and weights
```{r}
X_des = cbind(intercept = 0, nhanes_df |> select(Age, weight))
num_strata = length(unique(nhanes_df$strata)) # total strata
L = 1440 # dimension of the functional domain
sampled_strata = 15
strata_in_nhanes = unique(nhanes_df$strata)
I_n = nrow(X_des)

## First stage: Select strata
dirichlet_probs = gtools::rdirichlet(1, rep(1, num_strata))
if(sampled_strata == num_strata) {
  selected_strata = strata_in_nhanes
} else {
  selected_strata = sample(
    strata_in_nhanes,
    size = sampled_strata,
    replace = FALSE,
    prob = dirichlet_probs
  ) %>% sort
}
p_strata_design <- as.vector(dirichlet_probs)

# For strata that were not selected, set their probability to zero.
p_strata_design[!(strata_in_nhanes %in% selected_strata)] <- 0
# Renormalize so that the selected strata probabilities sum to 1.
p_strata_design <- p_strata_design / sum(p_strata_design)
  # strata_in_nhanes

## Second-stage: Individual Sampling within each strata
final_sample_id <- c()
p_i_given_strata <- rep(NA, I) # probabilities of selecting each individual within strata
p_i_strata <- rep(NA, I)
X_des[,2] = scale(X_des[,2])

for (strata in selected_strata) {
  # subjects in the specified strata
  inds <- which(nhanes_df$strata == strata)
  
  # Calculate inclusion probs
  inclusion_probs <- exp( (X_des[inds, 3] + X_des[inds, 2]) / 4)
    # Compute individual inclusion probabilities based on Age, dividing by 2 to reduce skewness
  inclusion_probs <- inclusion_probs / sum(inclusion_probs)
    # normalize
  inclusion_probs <- inclusion_probs * I_n * 2
    # Scale to get the desired number of individuals in ech strata / PSU combo
  
  inclusion_probs[inclusion_probs > 1] <- 1
  p_i_given_strata[inds] <- inclusion_probs
  
  # sample individuals based on inclusion probabilities
  sampled_in_strata <- inds[rbinom(length(inds), 1, inclusion_probs) == 1]
  final_sample_id <- c(final_sample_id, sampled_in_strata) # add indices to final sample
}

nhanes_df[final_sample_id,]

## Generate new weights from selection probs
strata_p_df = cbind(strata = strata_in_nhanes,
                    p_strata = p_strata_design)

df_add_w = cbind(nhanes_df, p_i_given_strata = p_i_given_strata) |>
  merge(strata_p_df, by = 'strata') |>
  mutate(weight_new = p_i_given_strata * p_strata)
```

```{r}
var_name = c('Intercept', 'Beta')
L = 1440
dat = read_rds('results/NI/NI_noOrigial_visual_only3_verify_raw.rds')
get_cis(betaTilde_boot = dat$res[[1]], 
        betaHat = dat$betaHat, 
        L=1440)

nknots_min = NULL
nknots_min_cov = 35
argvals = 1:L
B = ncol(betaTilde_boot[1, , ])
nknots <- min(round(L / 2), nknots_min)
nknots_cov <- ifelse(is.null(nknots_min_cov), 35, nknots_min_cov)
nknots_fpca <- min(round(L / 2), 35)
betaHat_boot <- array(NA, dim = c(nrow(betaHat), 
                                  ncol(betaHat), 
                                  ncol(betaTilde_boot[1, , ])))
betaHat.var <- array(NA, dim = c(L, L, nrow(betaHat)))
  # smooth

for (b in 1:B) {
  betaHat_boot[, , b] <- t(apply(betaTilde_boot[, , b], 1, function(x)
    gam(x ~ s(
      argvals, bs = "tp", k = (nknots + 1)
    ), method = "GCV.Cp")$fitted.values))
} # smooth each bootstrapped estimate (functional domain)
  
  
for (r in 1:nrow(betaHat)) {
  if (smooth_for_variance) {
    betaHat.var[, , r] <- 1.2 * var(t(betaHat_boot[r, , ]))
  } else{
    betaHat.var[, , r] <- 1.2 * var(t(betaTilde_boot[r, , ]))
  }
}
N <- 10000
qn <- rep(0, length = nrow(betaHat))
set.seed(456)
for (i in 1:length(qn)) {
  if (smooth_for_ci) {
    est_bs <- t(betaHat_boot[i, , ])
  } else{
    est_bs <- t(betaTilde_boot[i, , ])
  }
  fit_fpca <- suppressWarnings(refund::fpca.face(est_bs, knots = nknots_fpca)) # suppress sqrt(Eigen$values) NaNs
  ## extract estimated eigenfunctions/eigenvalues
  phi <- fit_fpca$efunctions
  lambda <- fit_fpca$evalues
  K <- length(fit_fpca$evalues)
  
  ## simulate random coefficients
  theta <- matrix(stats::rnorm(N * K), nrow = N, ncol = K) # generate independent standard normals
  if (K == 1) {
    theta <- theta * sqrt(lambda) # scale to have appropriate variance
    X_new <- tcrossprod(theta, phi) # simulate new functions
  } else{
    theta <- theta %*% diag(sqrt(lambda)) # scale to have appropriate variance
    X_new <- tcrossprod(theta, phi) # simulate new functions
  }
  x_sample <- X_new + t(fit_fpca$mu %o% rep(1, N)) # add back in the mean function
  Sigma_sd <- Rfast::colVars(x_sample, std = TRUE, na.rm = FALSE) # standard deviation: apply(x_sample, 2, sd)
  x_mean <- colMeans(est_bs)
  un <- rep(NA, N)
  for (j in 1:N) {
    un[j] <- max(abs((x_sample[j, ] - x_mean) / Sigma_sd))
  }
  qn[i] <- stats::quantile(un, 0.95)
}

  return(list(
    betaHat = betaHat,
    betaHat.var = betaHat.var,
    qn = qn
  ))






















results.lr = read_rds('data/nhanes_true_effect_lr.rds')
mod_output = dat[[1]]
beta_true = t(results.lr |> select(-Variable))[, 1:L]
cover_joint <- logical(nrow(beta_true))  # Using logical instead of NA-filled vector

source('src/00_sim_data_function.R')
source('src/01_sim_functions.R')

sqrt_diag_var <- apply(mod_output$betaHat.var, 3, function(mat) sqrt(diag(mat)))

results = data.frame()
for (p in 1:length(var_name)){
  # Extract beta and SE estimate
  true <- beta_true[p,]
  sqrt_diag_p <- sqrt_diag_var[, p]  # Extract precomputed diagonal elements
  
  # Compute upper and lower bounds in one step
  ## joint-wise
  margin <- mod_output$qn[p] * sqrt_diag_p #joint
  upper <- mod_output$betaHat[p,] + margin
  lower <- mod_output$betaHat[p,] - margin
  
  ## point-wise
  margin_pw <- 1.96 * sqrt_diag_p
  upper_pw <- mod_output$betaHat[p,] + margin_pw
  lower_pw <- mod_output$betaHat[p,] - margin_pw
  
  # return data for plots
  tmp = data.frame(type = var_name[p],
             beta_estimate = mod_output$betaHat[p,],
             beta_true = true,
             beta_pw_lower = lower_pw,
             beta_pw_upper = upper_pw,
             beta_jw_lower = lower,
             beta_jw_upper = upper)
  
  results = rbind(results, tmp)
}

get_coverage_stats(mod_output, beta_true, L = 1440, name = 'Preston')

a = results |>
  mutate(x = rep(seq(1:L), 2)) |>
  ggplot(aes(x = x)) +
  
  ## joint-wise CI
  geom_ribbon(aes(ymin = beta_pw_lower, ymax = beta_pw_upper), fill = "blue", alpha = 0.3) +
  ## point-wise CI
  geom_ribbon(aes(ymin = beta_jw_lower, ymax = beta_jw_upper), fill = "red", alpha = 0.3) +
  ## true values
  geom_line(aes(y = beta_true, color = "True value"), size = 0.5) +
  ## estimated values
  geom_line(aes(y = beta_estimate, color = "Estimated value"), size = 0.5) +
  
  ## Add non-covered points
  geom_point(data = results |> 
               mutate(x = rep(seq(1:L), 2)) |> 
               mutate(cover_jw = ifelse(beta_true >= beta_jw_lower & beta_true <= beta_jw_upper, 1, 0)) |> 
               filter(cover_jw == 0),
             aes(x = x, y = beta_true), 
             shape = 1, size = 2, color = "green") +
  facet_wrap(~type, scales = "free_y") + 
  labs(x = "X", y = "Y", title = "True vs Estimated with Confidence Intervals", color = "Line Type") +
  
  scale_color_manual(values = c("True value" = "black", "Estimated value" = "blue"))

a

## too smooth, add knots?

```


# 02. Informative Sampling

```{r}
nhanes_df |> filter(strata == 91)
```
