# 00. Prepare Data

## Load Packages

```{r}
library(dplyr)
library(knitr)
library(readxl)
library(readr)
library(haven)
library(visdat)
library(sjstats)
library(openxlsx)
library(stargazer)
library(broom)
library(ggplot2)
library(sjPlot)
library(stringr)
library(caret)
library(future)
library(purrr)
library(furrr)
library(here)
library(svrep)
library(patchwork)
```

## Load Data (single level)
In the current setting, we view it as the whole population, and the original survey design would not be considered.
```{r}
nhanes_df = readRDS('data/pa_df_persub.rds') |> 
  dplyr::rename(Age = age_in_years_at_screening) |>
  dplyr::rename(weight = full_sample_2_year_mec_exam_weight)

## subjects in the sample
length(unique(nhanes_df$SEQN)) # N = 13491
head(nhanes_df)
```

## Select population samples

For now, we didn't set any variable restrictions (like Age and Sex)

## Estimate True Population Effect

$$
y_i(s) = \beta_0(s) + X_i \beta_1(s) + \epsilon_i(s), \quad s = 1, 2, \ldots, 1440
$$

-   $y_i(s)$: PA monitoring data for individual i

-   $X_i$: Scalar covariate, and we consider age here

-   $\beta_0(s)$: Time-varying regression coefficients

-   $\epsilon_i(s) \sim N(0, \sigma^2(s))$

We need find a way to estimate true coefficients, $\beta_{1(s)}$ and intercepts $\beta_{0(s)}$. Here, we consider the same way in fast univariate inference (FUI) to estimate point-wise effect. i.e. use generalized survey-weighted linear mixed effects model in each time point throughout the functional domain.

In the current scenario, we consider the simpler situation that single level and survey weighted generalized linear regression (svyglm).

```{r}
# Estimate point-wise effect
## Response Variable
min_columns <- nhanes_df %>% select(starts_with("min")) %>% names()
```

### Linear regression model

```{r}
## Obtain True Effects: Regression at each time point
# results.lr <- map_dfr(min_columns, ~ {
#   model <- lm(reformulate("Age", response = .x), data = nhanes_df)
#   tidy(model) %>%
#     filter(term == "Age") %>%
#     mutate(Variable = .x, Intercept = coef(model)[1], Beta = estimate) %>%
#     select(Variable, Intercept, Beta)
# })
# write_rds(results.lr, 'data/nhanes_true_effect_lr.rds')


## Smooth estimated effects
results.lr = read_rds('data/nhanes_true_effect_lr.rds')
intercept_smooth <- smooth.spline(results.lr$Intercept)
beta_smooth <- smooth.spline(results.lr$Beta)
  # Here, we use GCV to select spar (cubic smoothing spline)
  # what method we should use for smoothing true estimates?
results.lr$Intercept_smoothed <- predict(intercept_smooth)$y
results.lr$Beta_smoothed <- predict(beta_smooth)$y


## True Coefficients and intercept (& Visualization)
results.lr |>
  mutate(Minutes = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable) |>
  tidyr::pivot_longer(cols = c(Beta, Intercept_smoothed, Intercept, Beta_smoothed), 
                      names_to = "Type",
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value)) +
  geom_point(color = "blue",
             alpha = 0.6, size = 0.8) +
  geom_smooth(method = "loess",
              linewidth = 0.8, 
              color = "red", se = TRUE, span = 0.3) +
  facet_wrap(~ Type, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Beta") +
  theme(text = element_text(size = 14))


## Export smoothed estimates
results.lr = results.lr |>
  select(-Intercept, -Beta) |>
  rename(Intercept = Intercept_smoothed,
         Beta = Beta_smoothed)
# write_rds(results.lr, 'data/nhanes_true_effect_lr_smoothed.rds')
```

### Survey weighted linear regression model

**WTINT2YR:** Full Sample 2 Year Interview Weight (full sample of participants who completed the NHANES household interview over a two-year survey period)

-   All participants who completed the household interview

**WTMEC2YR:** Full Sample 2 Year MEC Exam Weight (full sample of participants who completed both the household interview and the Mobile Examination Center (MEC) examination over a two-year survey period)

-   Only participants who completed the MEC examination (Lab Measurements)

```{r}
library(survey)
## Regression at each time point
head(nhanes_df)
design <- svydesign(ids = ~1, 
                    weights = ~weight, 
                    strata = ~strata,
                    psu = ~psu,
                    data = nhanes_df)

# results.svylr <- map_dfr(min_columns, ~ {
#   model <- svyglm(reformulate("Age", response = .x),
#                   design = design)
#   tidy(model) %>%
#     filter(term == "Age") %>%
#     mutate(Variable = .x, Intercept = coef(model)[1], Beta = estimate) %>%
#     select(Variable, Intercept, Beta)
# })
results.svylr = read_rds('data/nhanes_true_effect_svylr.rds')

## True Coefficients and intercept (& Visualization)
results.svylr |>
  mutate(Minutes = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable) |>
  tidyr::pivot_longer(cols = c(Beta, Intercept), 
                      names_to = "Type",
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value)) +
  geom_point(color = "blue",
             alpha = 0.6, size = 0.8) +       # 绘制数据点
  geom_smooth(method = "loess",
              linewidth = 0.8, 
              color = "red", se = TRUE, span = 0.3) +
  facet_wrap(~ Type, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Beta") +
  theme(text = element_text(size = 14))
```

### Save true effects

```{r}
# write_rds(results.svylr, 'data/nhanes_true_effect_svylr.rds')
# write_rds(results.lr, 'data/nhanes_true_effect_lr.rds')
```

#### Compare true effects from svylm and lm

```{r}
results.svylr = read_rds('data/nhanes_true_effect_svylr.rds')
results.lr = read_rds('data/nhanes_true_effect_lr.rds')

df = merge(results.svylr, 
      results.lr, 
      by = 'Variable',
      suffixes = c("_svylr","_lr")) |>
  mutate(Minute = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable)

df |>
  select(Intercept_svylr, Beta_svylr, Minute) |>
  mutate(type = "svylr") |>
  rename(Intercept = Intercept_svylr, Beta = Beta_svylr) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minute) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minute, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

```{r}
source('src/00_sim_data_function.R')
source('src/01_sim_functions.R')
```

# 01. Non-Informative Sampling
## Ignore original survey setting

1.  Generate strata selection probabilities ($H=30$) from Dirichlet distribution
2.  Sample $N=\{10, 20, 30\}$ strata in the simulation
3.  For each strata $h$ in the selected strata:
    -   Standardize Age into mean=0 and SD=1
    -   Calculate inclusion probabilities $p_i = exp(Age_{i}/2)$
    -   Normalize: $p_{i'} = \frac{p_i}{\sum_{i \in h}p_i}$
    -   Adjust by expected number of individuals sampled $p_{i''} = min(p_{i'} * I_n * 2, 1)$
    -   Obtain sample using binomial sampling

### Simulate Data
```{r}
results.lr.tmp = read_rds('data/nhanes_true_effect_lr_smoothed.rds') |>
  mutate(Minutes = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable)

X_des = cbind(intercept = 0, nhanes_df |> select(Age))
num_strata = 30 # total strata
L = 150 # dimension of the functional domain
sampled_strata = 15
Y_obs = nhanes_df[,which(grepl('min_', colnames(nhanes_df)))[1:L] ]

sim.data = simulate_nhanes_survey_fosr(X_des, 
                                       Y_obs,
                                       L = L,
                                       sampled_strata = 15)
```

### Survey design/weighted method
#### Visualize Point-Estimate
```{r}
betaTilde = get_betatilde(sim.data,
              family = "gaussian", type = "design",
              model_formula = as.formula('Y~X'))
betaHat = get_betahat(betaTilde, L = L)

## Visualize differences
df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```


#### Visualize 95%CI
```{r}
# betaTilde_boot = run_boots(
#   sim.data,
#   boot_type = 'Preston',
#   betaHat = betaHat,
#   L = L,
#   num_boots = 300,
#   model_formula = as.formula('Y~X')
# )
# 
# mod_output = get_cis(betaTilde_boot,
#         betaHat,
#         smooth_for_ci = TRUE,
#         smooth_for_variance = FALSE, # False
#         L = L)
# 
# results = get_coverage_stats(mod_output,
#                              beta_true = t(results.lr |> select(-Variable))[,1:L],
#                              name = 'Preston',
#                              # boot_type
#                              L = L)
# 
# write_rds(betaHat, 'tmp_result/betaHat_NI_noOrigial_survey_betaHat.rds')
# write_rds(betaTilde_boot, 'tmp_result/betaHat_NI_noOrigial_survey_betaTilde_boot.rds')
# write_rds(mod_output, 'tmp_result/betaHat_NI_noOrigial_survey_mod_output.rds')
```




### No survey design/Unweighted
#### Visualize Point-Estimate
```{r}
betaTilde = get_betatilde(sim.data,
              family = "gaussian", type = "none",
              model_formula = as.formula('Y~X'))
betaHat = get_betahat(betaTilde, L = L)

df = as.data.frame(t(betaTilde)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

#### Visualize 95%CI
```{r}
# betaTilde_boot = run_boots(
#   sim.data,
#   boot_type = 'Preston',
#   betaHat = betaHat,
#   L = L,
#   num_boots = 300,
#   model_formula = as.formula('Y~X')
# )
# 
# mod_output = get_cis(betaTilde_boot,
#         betaHat,
#         smooth_for_ci = TRUE,
#         smooth_for_variance = FALSE, # False
#         L = L)
# 
# results = get_coverage_stats(mod_output,
#                              beta_true = t(results.lr |> select(-Variable))[,1:L],
#                              name = 'Preston',
#                              # boot_type
#                              L = L)
# 
# write_rds(betaHat, 'tmp_result/betaHat_NI_noOrigial_NOsurvey_betaHat.rds')
# write_rds(betaTilde_boot, 'tmp_result/betaHat_NI_noOrigial_NOsurvey_betaTilde_boot.rds')
# write_rds(mod_output, 'tmp_result/betaHat_NI_noOrigial_NOsurvey_mod_output.rds')
```



### Simulation results (N=200) 
#### Merge results
```{r}
L = 1440
all_results <- list.files("results/NI/NI_noWeight_FPCA_noSmooth_adjustSTD/", 
                          pattern = "NI_noWeight_FPCA_noSmooth_adjustSTD.*\\.rds", 
                          full.names = TRUE) %>%
  map_dfr(readRDS)

summary_design = all_results %>%
  group_by(boot_type, var) %>%
  summarise(
    MISE = mean(MISE),
    mean_joint_se = mean(mean_joint_se),
    mean_pw_se = mean(mean_pw_se),
    cover_joint_ratio = mean(cover_joint),
    cover_pw_ratio = mean(cover_pw) / L,
    .groups = 'drop'
  )

formatted_table <- summary_design %>%
  mutate(across(where(is.numeric), ~round(., 4)))  # 四舍五入到4位小数

kable(summary_design,
      digits = 4,  # control digital number 
      caption = "Summary Statistics",  # Simple Caption
      format = "html")

## In the current setting, the length of domain will affect joint-wise coverage ratio.
## How about changing knot points?
## Test it.
```

## Consider original weights
### Simulate data
```{r}
X_des = cbind(intercept = 0, nhanes_df |> select(Age, weight))
X_des$weight = scale(X_des$weight)
num_strata = 30 # total strata
L = 150 # dimension of the functional domain
sampled_strata = 15
Y_obs = nhanes_df[,which(grepl('min_', colnames(nhanes_df)))[1:L] ]

sim.data = simulate_nhanes_survey_fosr_w(X_des, Y_obs,
                                         L = L,
                                         sampled_strata = 15)
```

### Survey design/weighted method
#### Visualize Point-Estimate
```{r}
betaTilde = get_betatilde(sim.data,
              family = "gaussian", type = "design",
              model_formula = as.formula('Y~X'))
betaHat = get_betahat(betaTilde, L = L)

## Visualize differences
df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) + 
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

#### Visualize 95%CI
```{r}
# betaTilde_boot = run_boots(
#   sim.data,
#   boot_type = 'Preston',
#   betaHat = betaHat,
#   L = L,
#   num_boots = 300,
#   model_formula = as.formula('Y~X')
# )
# 
# mod_output = get_cis(betaTilde_boot,
#         betaHat,
#         smooth_for_ci = TRUE,
#         smooth_for_variance = FALSE, # False
#         L = L)
# 
# results = get_coverage_stats(mod_output,
#                              beta_true = t(results.lr |> select(-Variable))[,1:L],
#                              name = 'Preston',
#                              # boot_type
#                              L = L)
# 
# write_rds(betaHat, 'tmp_result/betaHat_NI_oW_survey_betaHat.rds')
# write_rds(betaTilde_boot, 'tmp_result/betaHat_NI_oW_survey_betaTilde_boot.rds')
# write_rds(mod_output, 'tmp_result/betaHat_NI_oW_survey_mod_output.rds')
```

#### Simulation multiple times
```{r}
# nsim = 50
# parallel = TRUE
# ncores = parallelly::availableCores() - 1
# L = 150
# B = 300
# family = "gaussian"
# # boot_types = c('Rao-Wu-Yue-Beaumont', 'BRR', 'Preston', 'weighted', 'none')
# # fit_types = c('design', 'design', 'design', 'weighted', 'none')
# 
# boot_types = c('Preston', 'weighed', 'none')
# fit_types = c('design', 'weighted', 'none')
# 
# sim_res <- list() ## store simulation results
# for (iter in 1:nsim) {
#   set.seed(iter)
#   x = try({
#     data <- simulate_nhanes_survey_fosr_w(X_des, Y_obs, L = L, sampled_strata = 15)
#     
#     betaTilde = map(.x = fit_types, .f = get_betatilde, data = data, family = family)
#     betaHat = map(.x = betaTilde, get_betahat, L = L)
#     if (parallel) {
#       plan(multisession, workers = ncores)
#       res = future_map2(
#         .x = boot_types,
#         .y = betaHat,
#         .f = run_boots,
#         data = data,
#         # family = family,
#         num_boots = B,
#         set_seed = TRUE,
#         seed = 2025,
#         L = L,
#         .options = furrr_options(seed = TRUE)
#       )
#       cis = future_map2(
#         .x = res,
#         .y = betaHat,
#         .f = get_cis,
#         L = L,
#         smooth_for_ci = FALSE,
#         .options = furrr_options(seed = TRUE)
#       )
#       plan(sequential)
#     } else{
#       res = map2(
#         .x = boot_types,
#         .y = betaHat,
#         .f = run_boots,
#         data = data,
#         # family = family,
#         num_boots = B,
#         set_seed = TRUE,
#         seed = 2025,
#         L = L)
#       cis = map2(
#         .x = res,
#         .y = betaHat,
#         .f = get_cis,
#         smooth_for_ci = TRUE)
#     }
#     stats = map2(
#       .x = cis,
#       .y = boot_types,
#       .f = get_coverage_stats,
#       L = L,
#       beta_true = t(results.lr |> select(-Variable))[,1:L]
#     ) %>%
#       list_rbind() %>%
#       mutate(n = nrow(data), n_boot = B)
#     sim_res[[iter]] <- stats
#   })
#   rm(x)
#   print(iter)
# }
# 
# result =
#   sim_res %>%
#   keep(., is.data.frame) %>%
#   list_rbind(names_to = "id")
# 
# write_rds(result, 'tmp_result/betaHat_NI_oW_survey_results.rds')
```


### No survey design/Unweighted
#### Visualize Point-Estimate
```{r}
betaTilde = get_betatilde(sim.data,
              family = "gaussian", type = "none",
              model_formula = as.formula('Y~X'))
betaHat = get_betahat(betaTilde, L = L)

## Visualize differences
df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) + 
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

#### Visualize 95%CI
```{r}
# betaTilde_boot = run_boots(
#   sim.data,
#   boot_type = 'Preston',
#   betaHat = betaHat,
#   L = L,
#   num_boots = 300,
#   model_formula = as.formula('Y~X')
# )
# 
# mod_output = get_cis(betaTilde_boot,
#         betaHat,
#         smooth_for_ci = TRUE,
#         smooth_for_variance = FALSE, # False
#         L = L)
# 
# results = get_coverage_stats(mod_output,
#                              beta_true = t(results.lr |> select(-Variable))[,1:L],
#                              name = 'Preston', # boot_type
#                              L = L)
# 
# write_rds(betaHat, 'tmp_result/betaHat_NI_oW_NOsurvey_betaHat.rds')
# write_rds(betaTilde_boot, 'tmp_result/betaHat_NI_oW_NOsurvey_betaTilde_boot.rds')
# write_rds(mod_output, 'tmp_result/betaHat_NI_oW_NOsurvey_mod_output.rds')
```

### Simulation results (N=200) 
#### Merge results
```{r}
L = 1440
all_results <- list.files("results/NI/NI_noWeight_noSmooth//", 
                          pattern = "NI_noWeight_noSmooth_.*\\.rds", 
                          full.names = TRUE) %>%
  map_dfr(readRDS)


summary_design = all_results %>%
  group_by(boot_type, var) %>%
  summarise(
    MISE = mean(MISE),
    mean_joint_se = mean(mean_joint_se),
    mean_pw_se = mean(mean_pw_se),
    cover_joint_ratio = mean(cover_joint),
    cover_pw_ratio = mean(cover_pw) / L,
    .groups = 'drop'
  )

formatted_table <- summary_design %>%
  mutate(across(where(is.numeric), ~round(., 4)))  # 四舍五入到4位小数

kable(summary_design,
      digits = 4,  # 控制小数位数
      caption = "Summary Statistics",  # 简单标题
      format = "html")

## 0422: increase knots for estimating covariance and smooth boostratped trajectory
## 0423: increase knots for fPCA, no use
## double: double the number of knots
## try different simulated number (1-200 ->1-400): 0.9075

## candidate steps: increase scaling factors
## knot doesn't matter after reaching certain values.
## visualize current issue
## try 


all_results |>
  filter(var == 'Beta') |>
  filter(boot_type == 'Preston') |>
  filter(cover_joint == FALSE)
```
