---
title: "Emprical Simulation"
format: 
  html:
    code-fold: true
execute:
  warning: false
  echo: true
bibliography: references.bib
---

```{r,warning=FALSE, results = 'hide', collapse=TRUE}
library(dplyr)
library(readxl)
library(readr)
library(knitr)
library(haven)
library(visdat)
library(sjstats)
library(openxlsx)
library(stargazer)
library(broom)
library(ggplot2)
library(sjPlot)
library(stringr)
library(caret)
library(future)
library(purrr)
library(furrr)
library(here)
library(svrep)
library(patchwork)
```

------------------------------------------------------------------------

# Estimate True Population Effect

## Select population samples

Now, we didn't set any selection criteria (like Age and Sex)

## Estimate True Population Effect

$$
y_i(s) = \beta_0(s) + X_i \beta_1(s) + \epsilon_i(s), \quad s = 1, 2, \ldots, 1440
$$

-   $y_i(s)$: Physical activity monitoring data for individual i

-   $X_i$: Scalar covariate, and we consider **age** here

-   $\beta_0(s)$: Time-varying regression coefficients

-   $\epsilon_i(s) \sim N(0, \sigma^2(s))$

We need to find a way to estimate the true coefficients, $\beta_{1(s)}$ and intercepts $\beta_{0(s)}$. For this purpose, we apply the same approach used in fast univariate inference (FUI) to estimate point-wise effects. Specifically, we employ linear regression / generalized survey-weighted linear mixed effects models at each time point across the functional domain.

Currently, we view the NHANES data as the super-population, so we could ignore the original survey design or consider partial design component in the process of selecting subsample.

When estimating true effects, we use linear regression (for single level) to estimate true effects of Age and true intercept at each time point throughout the day.

### Linear regression (point-estimate)
#### Estimate true effects using LR
```{r}
## Obtain True Effects: Regression at each time point
# results.lr <- map_dfr(min_columns, ~ {
#   model <- lm(reformulate("Age", response = .x), data = nhanes_df)
#   tidy(model) %>%
#     filter(term == "Age") %>%
#     mutate(Variable = .x, Intercept = coef(model)[1], Beta = estimate) %>%
#     select(Variable, Intercept, Beta)
# })
# write_rds(results.lr, 'data/nhanes_true_effect_lr.rds')


## Smooth estimated effects
results.lr = read_rds('data/nhanes_true_effect_lr.rds')
intercept_smooth <- smooth.spline(results.lr$Intercept)
beta_smooth <- smooth.spline(results.lr$Beta)
  # Here, we use GCV to select spar (cubic smoothing spline)
  # what method we should use for smoothing true estimates?
results.lr$Intercept_smoothed <- predict(intercept_smooth)$y
results.lr$Beta_smoothed <- predict(beta_smooth)$y


## Export smoothed estimates
results.lr = results.lr |>
  select(-Intercept, -Beta) |>
  rename(Intercept = Intercept_smoothed,
         Beta = Beta_smoothed)
# write_rds(results.lr, 'data/nhanes_true_effect_lr_smoothed.rds')
```


```{r,warning=FALSE, results = 'hide', collapse=TRUE}
results.lr = read_rds('data/nhanes_true_effect_lr_smoothed.rds')

## True Coefficients and intercept (& Visualization)
results.lr |>
  mutate(Minutes = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable) |>
  tidyr::pivot_longer(cols = c(Beta, Intercept), 
                      names_to = "Type",
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value)) +
  geom_point(color = "blue",
             alpha = 0.6, size = 0.8) +       # 绘制数据点
  geom_smooth(method = "loess",
              linewidth = 0.8, 
              color = "red", se = TRUE, span = 0.3) +
  facet_wrap(~ Type, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throughout the day",
       x = "Minutes",
       y = "Beta") +
  theme(text = element_text(size = 14))
```

# Model Performance

For the single level, we consider two scenarios (This process is the same as the simulation study):

-   \(1\) non-informative sampling

-   \(2\) informative sampling

For each scenario, we would explore should we consider the original survey design in the subsample selection process. We consider three conditions:

-   \(1\) No original design considered

-   \(2\) Consider original weights

-   \(3\) Consider original strata and weights (TBD).

Finding: Condition (2) outperforms condition (1) in visualization quality, yet both conditions exhibit comparable performance across evaluation metrics.

## 01. Non-Informative Sampling

### Ignore original survey design in the selection process

1.  Generate strata selection probabilities ($H=30$) from Dirichlet distribution
2.  Sample $N=\{15\}$ strata in the simulation
3.  For each strata $h$ in the selected strata:
    -   Standardize Age into mean=0 and SD=1
    -   Calculate inclusion probabilities $p_i = exp(Age_{i}/2)$
    -   Normalize: $p_{i'} = \frac{p_i}{\sum_{i \in h}p_i}$
    -   Adjust by expected number of individuals sampled $p_{i''} = min(p_{i'} * I_n * 2, 1)$
    -   Obtain sample using binomial sampling

-   Only show first 150 time points here due to running time \[Need update after finish running\]

#### Not consider current design
```{r}
results.lr.tmp = read_rds('data/nhanes_true_effect_lr_smoothed.rds') |>
  mutate(Minutes = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable)
betaHat = readRDS('tmp_result/betaHat_NI_noOrigial_NOsurvey_betaHat.rds')
df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

#### Consider current design
```{r}
results.lr.tmp = read_rds('data/nhanes_true_effect_lr_smoothed.rds') |>
  mutate(Minutes = as.numeric(sub('min_', '', Variable)) ) |>
  select(-Variable)
betaHat = readRDS('tmp_result/betaHat_NI_noOrigial_survey_betaHat.rds')

df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```

#### Simulation (N=50)
We repeat the subsampling above and model fitting procedure 50 times.

*Summary Table*
```{r}
result_design = read_rds('tmp_result/betaHat_NI_noOrigial_survey_results.rds') 
  # Consider current design

summary_design = result_design %>%
  group_by(boot_type, var) %>%
  summarise(
    MISE = mean(MISE),
    mean_joint_se = mean(mean_joint_se),
    mean_pw_se = mean(mean_pw_se),
    cover_joint_ratio = mean(cover_joint),
    cover_pw_ratio = mean(cover_pw) / 150,
    .groups = 'drop'
  )

formatted_table <- summary_design %>%
  mutate(across(where(is.numeric), ~round(., 4)))

kable(summary_design,
      digits = 4,
      caption = "Summary Statistics", 
      format = "html")
```


### Consider original weights in the selection process
It didn't consider PA information here, but it's also a kind of infomrative manner to select subjects. [Q: Do we define it as informative sampling?]

1.  Generate strata selection probabilities ($H=30$) from Dirichlet distribution
2.  Sample $N=\{15\}$ strata in the simulation
3.  For each strata $h$ in the selected strata:
    -   Standardize *Age* and *Weight* into mean=0 and SD=1
    -   Calculate inclusion probabilities $p_i = exp( (Age_{i}+weight_{i}) /4)$
    -   Normalize: $p_{i'} = \frac{p_i}{\sum_{i \in h}p_i}$
    -   Adjust by expected number of individuals sampled $p_{i''} = min(p_{i'} * I_n * 2, 1)$
    -   Obtain sample using binomial sampling
    
#### Not consider current design
```{r}
betaHat = readRDS('tmp_result/betaHat_NI_oW_NOsurvey_betaHat.rds')
df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```


#### Consider current design
```{r}
betaHat = read_rds('tmp_result/betaHat_NI_oW_survey_betaHat.rds')
df = as.data.frame(t(betaHat)) |>
  rename(Beta = X, Intercept = `(Intercept)`) |>
  mutate(Minutes = row_number() ) |>
  merge(results.lr.tmp,
        by = 'Minutes',
        suffixes = c("_estimate","_lr"))

df |>
  select(Intercept_estimate, Beta_estimate, Minutes) |>
  mutate(type = "estimate") |>
  rename(Intercept = Intercept_estimate, Beta = Beta_estimate) |>
  bind_rows(
    df |>
      select(Intercept_lr, Beta_lr, Minutes) |>
      mutate(type = "lr") |>
      rename(Intercept = Intercept_lr, Beta = Beta_lr)
  ) |>
  tidyr::pivot_longer(cols = c(Intercept, Beta), 
                      names_to = "Metric", 
                      values_to = "Value") |>
  ggplot(aes(x = Minutes, y = Value, color = type)) +
  geom_point(alpha = 0.6, size = 0.5) +       # 绘制数据点
  facet_wrap(~ Metric, scales = "free_y") + 
  labs(title = "True effects of beta and intercept throught the day",
       x = "Minutes",
       y = "Value") +
  theme(text = element_text(size = 14))
```
#### Simulation (N=50)
We repeat the subsampling above and model fitting procedure 50 times.

*Summary Table*
```{r}
result_design = read_rds('tmp_result/betaHat_NI_oW_survey_results.rds') 
result_design_2 = read_rds('tmp_result/betaHat_NI_oW_survey_results_2.rds') 

  # Consider current design

summary_design = rbind(result_design, result_design_2) %>%
  group_by(boot_type, var) %>%
  summarise(
    MISE = mean(MISE),
    mean_joint_se = mean(mean_joint_se),
    mean_pw_se = mean(mean_pw_se),
    cover_joint_ratio = mean(cover_joint),
    cover_pw_ratio = mean(cover_pw) / 150,
    .groups = 'drop'
  )

formatted_table <- summary_design %>%
  mutate(across(where(is.numeric), ~round(., 4)))  # 四舍五入到4位小数

kable(summary_design,
      digits = 4,  # 控制小数位数
      caption = "Summary Statistics",  # 简单标题
      format = "html")
```

## 02. Informative Sampling (TBD)
Consider eigenvalues $\hat\lambda_k$in the selection process

