---
title: "Functional SL"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    code-background: true
    code-tools: true
    code-fold: true
    code-block-border-left: true
execute:
  echo: true
  cache: true
  message: false
  warning: false
editor: source
---

```{r load packages}
#| eval: true
#| include: false
library(here)
library(devtools)
library(fastFMM)
library(haven)
library(dplyr)
library(survey)
library(progress)
library(lme4)
library(mgcv)
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(tidyfun)
library(mvtnorm)
library(refund)
library(svrep)

```

# Generate data 

```{r}
#| echo: true
#| eval: false 


# generate Population Data
I <- 10000 ## number of subjects
L <- 50 ## dimension of the functional domain
SNR_sigma = 1
family <- "gaussian" ## family of longitudinal functional data

## specify true fixed effects functions
grid <- seq(0, 1, length = L)
beta_true <- matrix(NA, 2, L) # time-dependent beta-true trajectory
scenario <- 1 ## indicate the scenario to use
if(scenario == 1){
  beta_true[1,] = -0.15 - 0.1*sin(2*grid*pi) - 0.1*cos(2*grid*pi)
  beta_true[2,] = dnorm(grid, .6, .15)/20
}else if(scenario == 2){
  beta_true[1,] <- 0.53 + 0.06*sin(3*grid*pi) - 0.03*cos(6.5*grid*pi)
  beta_true[2,] <- dnorm(grid, 0.2, .1)/60 + dnorm(grid, 0.35, .1)/200 -
    dnorm(grid, 0.65, .06)/250 + dnorm(grid, 1, .07)/60
}
rownames(beta_true) <- c("Intercept", "x")


seed = 200
options(survey.lonely.psu = "adjust")

simulate_survey_fosr <- function(I = 10e6, # total population,
                                 I_n = 50, # subjects in each psu-strata combination
                                 alpha = 2, # number PSU
                                 L = 50, # dimension of the functional domain
                                 SNR_sigma = 1,
                                 num_strata = 30 # total strata
){
  # Generate X variable
  X_des = cbind(1, rnorm(I, 0, 2))  # design matrix: Intercept + predicted variables

  # Generate strata sizes using a Dirichlet distribution
  dirichlet_probs = gtools::rdirichlet(1, rep(1, num_strata))
  set.seed(seed)
  stratum_assignments = sample(1:num_strata, size = I, replace = TRUE, prob = dirichlet_probs)
  p_strata = table(stratum_assignments) / sum(table(stratum_assignments))
  p_strata = c(p_strata) %>% unname()
  w_strata = 1 / p_strata


  # Second-stag: Individual Sampling within each strata
  final_sample <- c()
  p_i_given_g <- rep(0, I)
  psu = c()

  for (strata in 1:num_strata) {
    inds <- which(stratum_assignments == strata)
    inclusion_probs <- exp(X_des[inds, 2])  # Compute individual inclusion probabilities based on covariate
    inclusion_probs <- inclusion_probs / sum(inclusion_probs) # normalize
    inclusion_probs <- inclusion_probs * I_n * 2 # Scale to get the desired number of individuals in each strata / PSU combo
    inclusion_probs[inclusion_probs > 1] <- 1
    sampled_in_strata <- inds[rbinom(length(inds), 1, inclusion_probs) == 1]
    final_sample <- c(final_sample, sampled_in_strata)
    p_i_given_g[inds] <- inclusion_probs
    psu_strata <- rbinom(length(sampled_in_strata), 1, 0.5) + 1
    psu <- c(psu, psu_strata)
  }
  length(final_sample) # 2939, close to 30 * 50 * 2

  # Compute final individual selection probabilities
  p_i <- p_strata[stratum_assignments] * p_i_given_g
  survey_weights <- 1 / p_i  # Final survey weights (Inf: Not selected)

  dat.sim <- data.frame(
    ID = final_sample,
    X = X_des[final_sample,2],
    strata = stratum_assignments[final_sample], # Assume same strata here ??
    psu = psu,
    weight = survey_weights[final_sample]
  )

  # Generate fixed effects and true Y.
  true_fixef <- X_des %*% beta_true
  true_fixef_sample <- true_fixef[final_sample,] # n x L matrix
  
  Y_obs <- matrix(NA, nrow(true_fixef_sample), L) 
  for(i in 1:nrow(true_fixef_sample)){
    for(j in 1:L){
        Y_obs[i, j] <- rnorm(1, mean = true_fixef_sample[i, j], sd = SNR_sigma / 5)
    }
  }
  
  Y_obs <- Y_obs + with(dat.sim, rnorm(alpha, sd = SNR_sigma)[psu])
  # Y_obs <- true_fixef_sample + with(dat.sim, rnorm(alpha, sd = SNR_sigma)[psu]) ## add some PSU-level variability in Y

  Y_obs <- data.frame(Y_obs)
  colnames(Y_obs) <- paste0("Y", 1:L)

  sample_data <- cbind(dat.sim, Y_obs)

  return(sample_data)
}
dat = simulate_survey_fosr()
write_rds(dat, here::here("sim_data", "dat_f2.rds"))
```

```{r}
data = read_rds(here::here("sim_data", "dat_f2.rds"))
```


## Explore data

We have $n = 2983$, function of length $L=50$ (outcome $Y$), predictor $X$, $30$ strata with two PSU per strata. 

```{r}
data %>% 
  glimpse()
```


### Coefficient functions (truth) 

```{r}
grid <- seq(0, 1, length = 50)

beta_true <- matrix(NA, 2, 50) # time-dependent beta-true trajectory
beta_true[1,] = -0.15 - 0.1*sin(2*grid*pi) - 0.1*cos(2*grid*pi)
beta_true[2,] = dnorm(grid, .6, .15)/20

rownames(beta_true) <- c("Intercept", "x")

beta_true %>% 
  as_tibble() %>% 
  mutate(beta = c("beta_0", "beta_1")) %>% 
  pivot_longer(cols = -beta) %>% 
  mutate(x = as.numeric(sub(".*V", "", name))) %>% 
  ggplot(aes(x = x, y = value)) + 
  facet_wrap(.~beta) + 
  geom_line() + 
  labs(x = "L", y = "Value", title = "True Coefficient Functions")

```

### Outcomes 

```{r}
Y_mat = data %>% select(starts_with("Y")) %>% 
  as.matrix()

data %>% 
  mutate(psu = paste0("PSU: ", psu)) %>% 
  select(-starts_with("Y")) %>% 
  mutate(Y = I(Y_mat)) %>% 
  mutate(Y = matrix(Y, ncol = 50),
         Y_tf = tfd(Y, length = 50),
         Y_smooth = tf_smooth(Y_tf, method = "lowess")) %>%
  group_by(psu) %>% 
  mutate(mean_curve = mean(Y_smooth)) %>%
  ungroup() %>% 
  slice(1:10) %>% 
  ggplot(aes(y = Y_smooth)) +
  geom_spaghetti(aes(y = mean_curve), lwd = 1.1) + 
  geom_spaghetti(lwd = .7, aes(color = factor(ID))) + 
  facet_wrap(.~psu) + 
  theme(legend.position = "none") +
  labs(x = "L", y = "Y (smoothed)", title = "Mean curves and subjects by PSU")

data %>% 
  mutate(psu = paste0("PSU: ", psu)) %>% 
  select(-starts_with("Y")) %>% 
  mutate(Y = I(Y_mat)) %>% 
  mutate(Y = matrix(Y, ncol = 50),
         Y_tf = tfd(Y, length = 50),
         Y_smooth = tf_smooth(Y_tf, method = "lowess")) %>%
  group_by(psu) %>% 
  mutate(mean_curve = mean(Y_smooth)) %>%
  ungroup() %>% 
  slice(1:10) %>% 
  ggplot(aes(y = Y_smooth)) +
  geom_spaghetti(aes(y = mean_curve), lwd = 1.1) +
  geom_spaghetti(lwd = .7, aes(color = factor(ID))) + 
  facet_wrap(.~psu) + 
  theme(legend.position = "none") +
  labs(x = "L", y = "Y (smoothed)", title = "Mean curves and subjects by PSU")

```



# FUI 

## Estimate $\tilde{\beta}$ 

```{r}
Y_mat <- data[, grepl('Y.', colnames(data))]
dat.fit <- data.frame(Y = Y_mat, data[, !grepl('Y.', colnames(data))])
formula.FUI = as.formula(paste0('Y~', 'X'))

weighted = TRUE
family = "gaussian"
unimm <- function(l){
    data$Yl <- unclass(data[,out_index][,l])

        svy_design <- svydesign(ids = ~ psu,
                            strata = ~ strata,
                            weights = ~ weight,
                            data = data,
                            nest = TRUE)
        # fit survey-weighted regression using svyglm
      if(weighted == TRUE){
        if(family == "gaussian"){
          fit_uni <- suppressMessages(svyglm(formula = stats::as.formula(paste0("Yl ~ ", model_formula[3])),
                                             design = svy_design,
                                             control = glm.control(maxit = 5000)))
        }else{
          fit_uni <- suppressMessages(svyglm(formula = stats::as.formula(paste0("Yl ~ ", model_formula[3])),
                                             design = svy_design,
                                             family = family,
                                             control = glm.control(maxit = 5000)))
        }} else{
        if(family == "gaussian"){
          fit_uni <- suppressMessages(lm(formula = stats::as.formula(paste0("Yl ~ ", model_formula[3])),
                                         data = data))
        }else{
          fit_uni <- suppressMessages(glm(formula = stats::as.formula(paste0("Yl ~ ", model_formula[3])),
                                          data = data,
                                          family = family) )
        }
      }

      ## fixed effects estimates
      if (weighted == TRUE) { ## this part seems redundant
        betaTilde <- coef(fit_uni)
      }else{
        betaTilde <- coef(fit_uni)
      }
   
      return(list(betaTilde = betaTilde))

}

model_formula <- formula.FUI
out_index <- grep(paste0("^", model_formula[2]), names(data)) # indices that start with the outcome name
if(length(out_index) != 1){ # functional observations stored in multiple columns
    L <- length(out_index)
  }else{ # observations stored as a matrix in one column using the I() function
    L <- ncol(data[,out_index])
  }
argvals <- 1:L

  
massmm <- lapply(argvals, unimm)

# Obtain betaTilde, fixed effects estimates
betaTilde <- t(do.call(rbind, lapply(massmm, '[[', 1)))
colnames(betaTilde) <- argvals

```


## Smooth

```{r}
L = 50
nknots_min = NULL
nknots_min_cov = 35
nknots <- min(round(L / 2), nknots_min)
nknots_cov <- ifelse(is.null(nknots_min_cov), 35, nknots_min_cov)
nknots_fpca <- min(round(L / 2), 35)

betaHat <- t(apply(betaTilde, 1, function(x) mgcv::gam(x ~ s(argvals, bs = "tp", k = (nknots + 1)), method = "GCV.Cp")$fitted.values))

rownames(betaHat) <- rownames(betaTilde)
colnames(betaHat) <- 1:L


```


```{r}
# plot 
bt_plot = 
  betaHat %>% 
    as_tibble() %>% 
    mutate(beta = c("beta_0", "beta_1")) %>% 
    pivot_longer(cols = -beta) %>% 
    mutate(x = as.numeric(sub(".*V", "", name))) 
btrue_plot = 
  beta_true %>% 
  as_tibble() %>% 
  mutate(beta = c("beta_0", "beta_1")) %>% 
  pivot_longer(cols = -beta) %>% 
  mutate(x = as.numeric(sub(".*V", "", name))) 

bt_plot %>% mutate(type = "estimate") %>% 
  bind_rows(btrue_plot %>% mutate(type = "true")) %>% 
  ggplot(aes(x = x, y = value, color = type)) + 
  facet_wrap(.~beta) + 
  geom_line() + 
  labs(x = "L", y = "Value", title = "Coefficient Functions")

```

## Inference 

### Create functions for bootstrapping 

```{r}
#| echo: true 

# function to generate betaTilde_boot: the bootstrapped replicates 
run_rwyb = function(data, num_boots, set_seed = TRUE, seed = 2025, betaHat, L = 50, save_reps = FALSE, nknots_min = NULL, nknots_min_cov = 35){
        B <- num_boots
        betaTilde_boot <- array(NA, dim = c(nrow(betaHat), ncol(betaHat), B))
        pb <- progress_bar$new(format = "  running across L [:bar] :percent eta: :eta",,
                               total = L)
        if(save_reps){
          reps = list()
        }
        for(l in 1:L){
          pb$tick()
          data$Yl <- unclass(data[,out_index][,argvals[l]])
          svy_design <- svydesign(ids = ~ psu,
                                  strata = ~ strata,
                                  weights = ~ weight,
                                  data = data,
                                  nest = TRUE)
          if(set_seed){
            set.seed(seed)
          } 
          rwyb_bootstrap <- svy_design %>% 
            as_bootstrap_design(
            type = "Rao-Wu-Yue-Beaumont", 
            replicates = B
          )

          coef.output = suppressWarnings(
            svyglm(formula = stats::as.formula(paste0("Yl ~ ", model_formula[3])),
                                                 design = rwyb_bootstrap,
                                                 return.replicates = TRUE,
                                                 control = glm.control(maxit = 5000)))

          betaTilde_boot[,l,] <- t(as.matrix(coef.output$replicates[,1:nrow(betaHat)]))
          if(save_reps){
            reps[[l]] <- coef.output$survey.design$repweights
          }
        }
        if(save_reps){
          return(
            list(
              betaTilde_boot = betaTilde_boot,
              bootstrap_samps = B,
              rep_list = reps
            )
          )
        } else{
          return(list(
            betaTilde_boot = betaTilde_boot,
            qn = qn,
            bootstrap_samps = B
          ))
        }
}

get_cb = function(betaTilde_boot, smooth_for_ci = TRUE, smooth_for_variance = TRUE, L = 50, betaHat, nknots_min = NULL, nknots_min_cov = 35, fpca_band = FALSE){
  B = ncol(betaTilde_boot[1,,])
  nknots <- min(round(L / 2), nknots_min)
  nknots_cov <- ifelse(is.null(nknots_min_cov), 35, nknots_min_cov)
  nknots_fpca <- min(round(L / 2), 35)
  betaHat_boot <- array(NA, dim = c(nrow(betaHat), ncol(betaHat), ncol(betaTilde_boot[1,,])))
  betaHat.var <- array(NA, dim = c(L, L, nrow(betaHat)))
  # smooth 
  for (b in 1:B) {
      betaHat_boot[, , b] <- t(apply(betaTilde_boot[, , b], 1, function(x)
        gam(x ~ s(
          argvals, bs = "tp", k = (nknots + 1)
        ), method = "GCV.Cp")$fitted.values))
    }
  for (r in 1:nrow(betaHat)) {
    if (smooth_for_variance) {
      betaHat.var[, , r] <- 1.2 * var(t(betaHat_boot[r, , ])) # scaling is correct??
    } else{
      betaHat.var[, , r] <- 1.2 * var(t(betaTilde_boot[r, , ])) # scaling is correct??
    }
  }

  N <- 10000
  qn <- rep(0, length = nrow(betaHat))
  if(fpca_band){
    set.seed(456)
    for(i in 1:length(qn)){
        if(smooth_for_ci){
          est_bs <- t(betaHat_boot[i,,])
        } else{
          est_bs <- t(betaTilde_boot[i,,])
        }
        fit_fpca <- suppressWarnings( refund::fpca.face(est_bs, knots = nknots_fpca)) # suppress sqrt(Eigen$values) NaNs
        ## extract estimated eigenfunctions/eigenvalues
        phi <- fit_fpca$efunctions
        lambda <- fit_fpca$evalues
        K <- length(fit_fpca$evalues)

        ## simulate random coefficients
        theta <- matrix(stats::rnorm(N*K), nrow=N, ncol=K) # generate independent standard normals
        if(K == 1){
          theta <- theta * sqrt(lambda) # scale to have appropriate variance
          X_new <- tcrossprod(theta, phi) # simulate new functions
        }else{
          theta <- theta %*% diag(sqrt(lambda)) # scale to have appropriate variance
          X_new <- tcrossprod(theta, phi) # simulate new functions
        }
        x_sample <- X_new + t(fit_fpca$mu %o% rep(1,N)) # add back in the mean function
        Sigma_sd <- Rfast::colVars(x_sample, std = TRUE, na.rm = FALSE) # standard deviation: apply(x_sample, 2, sd)
        x_mean <- colMeans(est_bs)
        un <- rep(NA, N)
        for(j in 1:N){
          un[j] <- max(abs((x_sample[j,] - x_mean)/Sigma_sd))
        }
        qn[i] <- stats::quantile(un, 0.95)
      }
  } else { 
    for(i in 1:length(qn)){
      if(smooth_for_ci){
        est_bs <- t(betaHat_boot[i,,])
      } else{
        est_bs <- t(betaTilde_boot[i,,])
      }
        Sigma <- var(est_bs)
        Cf = cov2cor(Sigma)
        qn[i] <- qmvnorm(1 -.05, corr = Cf, tail= "both.tails")$quantile
        # other way 
        # sample <- rmvnorm(N, mean = colMeans(est_bs), sigma = Sigma) 
        # get absolute value of sample
        # sample_abs <- abs(sample) # get abs
        # max_sample <- apply(sample_abs, 1, max) # get max 
        # q_alpha = stats::quantile(max_sample, 0.95) # get quantile 
      }
  }
  return(list(betaHat = betaHat, betaHat.var = betaHat.var, qn = qn))
}

```


```{r}
# plot betahat boot and beta tilde boot
B = 100
betaHat_boot <- betaTilde_boot <- betaHat_fpcaboot <- array(NA, dim = c(nrow(betaHat), ncol(betaHat), B))
for (l in 1:L) {
  data$Yl <- unclass(data[, out_index][, argvals[l]])
  svy_design <- svydesign(
    ids = ~ psu,
    strata = ~ strata,
    weights = ~ weight,
    data = data,
    nest = TRUE
  )
  
  set.seed(2024)
  
  rwyb_bootstrap <- svy_design %>%
    as_bootstrap_design(type = "Rao-Wu-Yue-Beaumont", replicates = B)
  
  coef.output = suppressWarnings(
    svyglm(
      formula = stats::as.formula(paste0("Yl ~ ", model_formula[3])),
      design = rwyb_bootstrap,
      return.replicates = TRUE,
      control = glm.control(maxit = 5000)
    )
  )
  
  betaTilde_boot[, l, ] <- t(as.matrix(coef.output$replicates[, 1:nrow(betaHat)]))
}  
  #smooth
for (b in 1:B) {
  betaHat_boot[, , b] <- t(apply(betaTilde_boot[, , b], 1, function(x)
    gam(x ~ s(
      argvals, bs = "tp", k = (25 + 1)
    ), method = "GCV.Cp")$fitted.values))
}

  
            
get_coefs = function(b){
  tibble(int_raw = betaTilde_boot[1,,b],
           int_smooth = betaHat_boot[1,,b],
           slope_raw = betaTilde_boot[2,,b],
           slope_smooth = betaHat_boot[2,,b],
           L = 1:50,
           rep = b)
}

plt_df = map_dfr(1:B, get_coefs)
plt_df %>% 
  filter(rep %in% sample(1:B, 10, replace = FALSE)) %>% 
  pivot_longer(cols = -c(L, rep)) %>% 
    separate_wider_delim(name, delim = "_", names = c("coef", "type"))  %>% 
    ggplot(aes(x = L, y = value)) + 
    facet_grid(type~coef) + 
    geom_line(aes(color = factor(rep))) + 
    scale_color_discrete(name = "Repetition")




```


```{r}
plot_res = function(sim_res, t = "", ylim = NULL) { 
  num_var = nrow(sim_res$betaHat)
  plot_df = 
    map_dfr(.x = 1:num_var,
      .f = function(r){
       data.frame(s = 1:length(sim_res$betaHat[r,]), 
                beta = sim_res$betaHat[r,], 
                lower = sim_res$betaHat[r, ] - 2 * sqrt(diag(sim_res$betaHat.var[, , r])), 
                upper = sim_res$betaHat[r, ] + 2 * sqrt(diag(sim_res$betaHat.var[, , r])), 
                lower.joint = sim_res$betaHat[r, ] - sim_res$qn[r] * 
                sqrt(diag(sim_res$betaHat.var[, , r])), 
                upper.joint = sim_res$betaHat[r, ] + sim_res$qn[r] * sqrt(diag(sim_res$betaHat.var[, , r]))) %>% 
          mutate(name = sim_res$betaHat %>% rownames() %>% .[r])
        
      })
  p = plot_df %>% 
    ggplot(aes(x = s, y = beta)) + 
    geom_line() + 
    geom_ribbon(aes(x = s, ymin = lower, ymax = upper), alpha = .3,
                fill = "gray10") + 
    geom_ribbon(aes(x = s, ymin = lower.joint, ymax = upper.joint), alpha = .3,
                fill = "gray20") + 
    facet_wrap(.~name) + 
    labs(x = "L", y = "Estimate", title = paste0(t)) + 
    theme_classic()
  if(!is.null(ylim)){
    p = p + scale_y_continuous(limits=ylim)
  }
  print(p)
}
```


### Run bootstrap 


$B = 500$, set seed within each $l$, and use FPCA method to get CMA bands. 


```{r}
#| eval: false 

s1_seed = run_rwyb(data = data,
                   num_boots = 500,
                   betaHat = betaHat, 
                   set_seed = TRUE,
                   save_reps = TRUE)

write_rds(s1_seed, here::here("sim_res", "rwyb_seed.rds"))
```

```{r}
s1_seed = read_rds(here::here("sim_res", "rwyb_seed.rds"))

## using fpca for CB 
svar_sci = get_cb(betaTilde_boot = s1_seed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = TRUE,
                   smooth_for_variance = TRUE,
                   fpca_band = TRUE)
svar_ci = get_cb(betaTilde_boot = s1_seed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = FALSE,
                   smooth_for_variance = TRUE,
                   fpca_band = TRUE)

var_sci = get_cb(betaTilde_boot = s1_seed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = TRUE,
                   smooth_for_variance = FALSE,
                   fpca_band = TRUE)

var_ci = get_cb(betaTilde_boot = s1_seed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = FALSE,
                   smooth_for_variance = FALSE,
                   fpca_band = TRUE)


```


```{r}
plot_res(svar_sci, t = "Smooth for Variance and for CB")
plot_res(svar_ci, t = "Smooth for Variance Only")
plot_res(var_sci, t = "Smooth for CB Only")
plot_res(var_ci, t = "No Smoothing")

```

Use multivariate normality to get CMA bands: 

```{r}
svar_sci_mvn = try({get_cb(betaTilde_boot = s1_seed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = TRUE,
                   smooth_for_variance = TRUE,
                   fpca_band = FALSE)})
svar_ci_mvn = get_cb(betaTilde_boot = s1_seed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = FALSE,
                   smooth_for_variance = TRUE,
                   fpca_band = FALSE)

var_sci_mvn = try({get_cb(betaTilde_boot = s1_seed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = TRUE,
                   smooth_for_variance = FALSE,
                   fpca_band = FALSE)})

var_ci_mvn = get_cb(betaTilde_boot = s1_seed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = FALSE,
                   smooth_for_variance = FALSE,
                   fpca_band = FALSE)

```

```{r}
plot_res(svar_ci_mvn, t = "Smooth for Variance Only")
plot_res(var_ci_mvn, t = "No Smoothing")

```

$B = 500$, do not set seed within each $l$, and use FPCA method to get CMA bands. 

```{r}
#| eval: false 
s1_noseed = run_rwyb(data = data,
                   num_boots = 500,
                   betaHat = betaHat, 
                   set_seed = FALSE,
                   save_reps = TRUE)

write_rds(s1_noseed, here::here("sim_res", "rwyb_noseed.rds"))
```

```{r}
s1_noseed = read_rds(here::here("sim_res", "rwyb_noseed.rds"))

## using fpca for CB 
ns_svar_sci = get_cb(betaTilde_boot = s1_noseed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = TRUE,
                   smooth_for_variance = TRUE,
                   fpca_band = TRUE)
ns_svar_ci = get_cb(betaTilde_boot = s1_noseed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = FALSE,
                   smooth_for_variance = TRUE,
                   fpca_band = TRUE)

ns_var_sci = get_cb(betaTilde_boot = s1_noseed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = TRUE,
                   smooth_for_variance = FALSE,
                   fpca_band = TRUE)

ns_var_ci = get_cb(betaTilde_boot = s1_noseed$betaTilde_boot,
                   betaHat = betaHat, 
                   smooth_for_ci = FALSE,
                   smooth_for_variance = FALSE,
                   fpca_band = TRUE)


```


```{r}
plot_res(ns_svar_sci, t = "Smooth for Variance and for CB")
plot_res(ns_svar_ci, t = "Smooth for Variance Only")
plot_res(ns_var_sci, t = "Smooth for CB Only")
plot_res(ns_var_ci, t = "No Smoothing")

```
# Weighting 


```{r}

svy_des =
  svydesign(
    ids = ~ psu,
    strata = ~ strata,
    weights = ~ weight,
    data = data,
    nest = TRUE
  )
svy_des 
rwyb_des = svy_des %>% 
  as_bootstrap_design(type = "Rao-Wu-Yue-Beaumont",
                      replicates = 500,
                      samp_method_by_stage = "PPSWOR")
# cluster should have two columns 
wts = make_rwyb_bootstrap_weights(svy_des,
                                  num_replicates = 500,
                                  samp_unit_ids = svy_des$cluster,
                                  samp_unit_sel_probs = svy_des$prob,
                                  strata_ids = svy_des$strata,
                                  samp_method_by_stage =  c("PPSWOR"))
rwyb_des2 <- 
  svrepdesign(
       data = svy_des$variables,
       repweights = wts,
       weights = weights(svy_des, type = "sampling"),
       type = "bootstrap"
     )
# try BRR 
brr_des = svy_des %>% 
  as.svrepdesign(type = "BRR")

jkn_des = svy_des %>% 
  as.svrepdesign(type = "JKn")

bs_des = svy_des %>% 
  as.svrepdesign(type = "bootstrap")


  ## compare 
svyglm(Y1 ~ X, design = svy_des) %>% broom::tidy() %>% pull(std.error) %>% .[2]
svyglm(Y1 ~ X, design = rwyb_des) %>% broom::tidy() %>% pull(std.error) %>% .[2]
svyglm(Y1 ~ X, design = rwyb_des2) %>% broom::tidy() %>% pull(std.error) %>% .[2]
svyglm(Y1 ~ X, design = brr_des) %>% broom::tidy() %>% pull(std.error) %>% .[2]
svyglm(Y1 ~ X, design = jkn_des) %>% broom::tidy() %>% pull(std.error) %>% .[2]
svyglm(Y1 ~ X, design = bs_des) %>% broom::tidy() %>% pull(std.error) %>% .[2]

```

```{r}
#| include: false 
#| eval: false 
data(api)
dclus1<-svydesign(id=~dnum, weights=~pw, data=apiclus1, fpc=~fpc)

rwyb_des = dclus1 %>% 
  as_bootstrap_design(type = "Rao-Wu-Yue-Beaumont",
                      replicates = 500)
multistage_srswor_design <- svydesign(data = mu284,
                                      ids = ~ id1 + id2,
                                      fpc = ~ n1 + n2)
des1 = as_bootstrap_design(multistage_srswor_design,
                           replicates = 500)

bootstrap_rep_design <- as_bootstrap_design(multistage_srswor_design,
                                            type = "Rao-Wu-Yue-Beaumont",
                                            replicates = 500)
bootstrap_replicate_weights <- make_rwyb_bootstrap_weights(
  num_replicates = 500,
  samp_unit_ids = multistage_srswor_design$cluster,
  strata_ids = multistage_srswor_design$strata,
  samp_unit_sel_probs = multistage_srswor_design$fpc$sampsize /
    multistage_srswor_design$fpc$popsize,
  samp_method_by_stage = c("SRSWOR", "SRSWOR")
)

     ## Create a replicate design object with the survey package
     bootstrap_rep_design2 <- svrepdesign(
       data = multistage_srswor_design$variables,
       repweights = bootstrap_replicate_weights,
       weights = weights(multistage_srswor_design, type = "sampling"),
       type = "bootstrap"
     )

bootstrap_rep_design <- as_bootstrap_design(multistage_srswor_design,
                                            type = "Rao-Wu-Yue-Beaumont",
                                            replicates = 500)
  

```
